---
title: "PSSN-II HFC and Variable Analysis"
output: html_document
date: "2024-07-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# Load necessary libraries
library(tidyverse)
library(haven)
library(lubridate)
library(dplyr)
library(tidyr)
library(ggplot2)

source("global_env_setup.R")
source("HFC_utils.R")
#Export HFC output in spreadsheet
#source("HFC_export_G_drive.R")
```

## Top 3 Items

```{r overall_completion}
hh_completion = read_dta(file.path(Sys.getenv("hfc_endline_data_dir"), "hh_completion.dta")) 
over_completion = mean(hh_completion$hh_status_completed)
```

- Overall completion rate is `r over_completion*100`%
- Supervisors with most outliers/enumerators: Enumerator7
- Key variable of interests with most outliers: asset_roof


## Completeness (Endline Pilot Data)
Overall completion rate: `r over_completion`

### Domain breakdown

```{r domain_completion}
domain_completion = read_dta(file.path(Sys.getenv("hfc_endline_data_dir"), "domain_completion.dta")) 

domain_completion %>% 
  arrange(desc(percent_compl_intw)) %>%
  ggplot(aes(x = domain, y = percent_compl_intw*100)) +
  geom_col(aes(fill=domain), width = 0.6) + coord_flip() +
  labs(title = "Survey completition rate by domain", x = "% Completed", y="Domain (Mainland vs Zanzibar)")
```

### Village completion
```{r village_completion}
library(knitr)
village_completion = read_dta(file.path(Sys.getenv("hfc_endline_data_dir"), "vlg_completion.dta")) %>%
  mutate(percent_compl_intw = nb_hh_completed/hh_target)


cols = c("village", "hh_target", "nb_hh_completed", "percent_compl_intw")


village_completion_table = village_completion %>%
  select(cols) %>%
  arrange(desc(percent_compl_intw))
  

kable(village_completion_table, format = "markdown", col.names = c("Village", "# of hh targeted", "# of hh completed", "Completion rate %"))

# 
# village_overview %>% 
#   arrange(percent_compl_intw) %>% 
#   head(5) %>%
#   ggplot(aes(x = village, y = percent_compl_intw)) +
#   geom_col(aes(fill=village), width = 0.6) + coord_flip() +
#   labs(title = "5 Villages with lowest completion rate", x = "Completion rate", y = "Village") 
```


### Enumerator breakdown
```{r enum_completion}

enum_completion = read_dta(file.path(Sys.getenv("hfc_endline_data_dir"), "enumerator_completion.dta")) %>%
  mutate(nb_total = nb_completed + nb_duplicates + nb_missingipv + nb_missingtrack) %>%
  mutate(percent_compl_intw = nb_completed/nb_total)


enum_completion %>%
  arrange(desc(percent_compl_intw)) %>%
  head(5) %>%
  ggplot(aes(x = enum_name, y = percent_compl_intw *100)) +
  geom_col(aes(fill=enum_name), width = 0.6) + coord_flip()
```

## Outliers/boxplots

```{r outliers, warning=FALSE, echo=FALSE}
source("HFC_outliers.R")
hh_plot_demo = hh_outlier_enhanced %>% filter(variable %in% c("main_resp_age", "head_age", "hh_size"))
ggplot(hh_plot_demo, aes(x = variable, y = value)) + geom_boxplot(aes(fill =
                                                                        variable))
hh_plot_assets = hh_outlier_enhanced %>% filter(variable %in% c("assets_roof", "assets_wall", "hhassets_12"))
ggplot(hh_plot_assets, aes(x = variable, y = value)) + geom_boxplot(aes(fill =
                                                                          variable))

hh_plot_consumption = hh_outlier_enhanced %>% filter(
  variable %in% c(
    "cereals_grain_amt",
    "electricity_cost",
    "spices_condmnt_amt",
    "tobacco_amt"
  )
)
ggplot(hh_plot_consumption, aes(x = variable, y = value)) + geom_boxplot(aes(fill = variable))


hh_plot_duration = hh_outlier_enhanced %>% filter(variable %in% c("duration"))
ggplot(hh_plot_duration, aes(x = variable, y = value)) + geom_boxplot(aes(fill = variable))
```


## Variables Analysis (Baseline Analysis Data)


### Demographics

```{r demographicis}
#household size, percentage of female-headed households
hh_constructed = read_dta(file.path(Sys.getenv("analysis_baseline_data_dir"), "HHSurvey.dta"))

hh_demo = hh_constructed %>%
  select(female_headed, tot_hhmbr) %>%
  mutate(hh_size = as.numeric(tot_hhmbr))

pct_female_hh = mean(hh_demo$female_headed, na.rm = TRUE)

hh_demo %>%
  na.omit(hh_size) %>%
  ggplot(aes(x = hh_size)) +
    geom_histogram(aes(y = ..density..), binwidth = 0.5, fill = "blue", color = "black", alpha = 0.7) +
    geom_density(color = "red", size = 1) +
    labs(title = "Household size distribution", x = "HH size", y = "Density") +
    theme_minimal()
```

% of Female household head HH is: `r pct_female_hh`


### Income and Savings

```{r income_savings}
#Income (e.g., non-farm revenue), savings account balance
hh_income = hh_constructed %>%
  select(starts_with("income"))

hh_income %>%
  na.omit(income_raw) %>%
  filter(income_raw>median(income_raw)-1.5*IQR(income_raw)&income_raw<median(income_raw)+1.5*IQR(income_raw)) %>%
  ggplot(aes(x = income_raw)) +
    geom_histogram(aes(y = ..density..), bins = 15, fill = "blue", color = "black", alpha = 0.7) +
    geom_density(color = "red", size = 1) +
    labs(title = "Household total income distribution", x = "HH total income (raw)", y = "Density") +
    theme_minimal()


#Savings
hh_savings = read_dta(file.path(Sys.getenv("analysis_baseline_data_dir"), "cs_roster2_savingaccount.dta"))

hh_savings %>%
  na.omit(cs_totalsaving) %>%
  filter(cs_totalsaving>median(cs_totalsaving)-1.5*IQR(cs_totalsaving)&cs_totalsaving<median(cs_totalsaving)+1.5*IQR(cs_totalsaving)) %>%
  ggplot(aes(x = cs_totalsaving)) +
    geom_histogram(aes(y = ..density..), bins = 15, fill = "blue", color = "black", alpha = 0.7) +
    geom_density(color = "red", size = 1) +
    labs(title = "Household total savings distribution", x = "HH total savings", y = "Density") +
    theme_minimal()



```

### Expenditures 
```{r expenditures}
#Expenditure on health, food, and non-food consumption

#Food
food_consumption_aggregate = read_dta(file.path(Sys.getenv("analysis_baseline_data_dir"), "consumption_food_exp_aggr.dta"))

food_consumption_aggregate %>%
  na.omit(tot_food_value) %>%
  filter(tot_food_value>median(tot_food_value)-1.5*IQR(tot_food_value)&tot_food_value<median(tot_food_value)+1.5*IQR(tot_food_value)) %>%
  ggplot(aes(x = tot_food_value)) +
    geom_histogram(aes(y = ..density..), bins = 15, fill = "blue", color = "black", alpha = 0.7) +
    geom_density(color = "red", size = 1) +
    labs(title = "Household total food consumption distribution", x = "HH total food consumption", y = "Density") +
    theme_minimal()

#Health consumption
health = hh_constructed %>%
  select(
    exp_health_year_raw,
    expenses_illness_year,
    expenses_prevention_year,
    expenses_medicine_year,
    expenses_overnight_year,
    expenses_traditional_year
  )

health %>%
  na.omit(exp_health_year_raw) %>%
  filter(exp_health_year_raw>0) %>%
  filter(exp_health_year_raw>median(exp_health_year_raw)-1.5*IQR(exp_health_year_raw)&exp_health_year_raw<median(exp_health_year_raw)+1.5*IQR(exp_health_year_raw)) %>%
  ggplot(aes(x = exp_health_year_raw)) +
    geom_histogram(aes(y = ..density..), bins = 15, fill = "blue", color = "black", alpha = 0.7) +
    geom_density(color = "red", size = 1) +
    labs(title = "Household total health expenditure distribution", x = "HH total health expenditure", y = "Density") +
    theme_minimal()

#Non-Farm consumption





```


### Participation (End line pilot data) 

```{r labor-participation}
#Dummy variables: PW participation, EL participation, Received transfers from PW, Received transfers from EL, Transfer modality: Electronic vs manual, Income generating activities: business ownership, livestock ownership, agriculture cultivation, labor income through private paid work (not PW)

hh_endline_data = read_dta(file.path(Sys.getenv("hfc_endline_data_dir"), "HHSurvey.dta"))

#write.csv(colnames(hh_endline_data), "hh_survey_cols.csv")

participation = hh_endline_data %>%
  select(
    pw_particip,
    el_particip,
    epay_particip
  )

participation_summary = data.frame(
  pw_particip_rate = mean(participation$pw_particip, na.rm=TRUE),
  el_particip_rate = mean(participation$el_particip, na.rm=TRUE),
  epay_particip_rate = mean(participation$epay_particip, na.rm=TRUE)
)

participation_summary <- participation_summary %>% 
  pivot_longer(
    everything(),
    names_to = "Participation",
    values_to = "Rate"
)

participation_summary %>%
  ggplot(aes(x = Participation, y = Rate *100)) +
  geom_col(aes(fill=Participation), width = 0.6) + coord_flip()

transfers = hh_endline_data %>%
  select(bg_received,
    cs_transfers,
    )
```

