---
title: "PSSN-II Baseline Analysis"
author: "Shaochen Huang"
date: "2024-08-19"
output: html_document
---
## Overview
This analysis explores the baseline data in complimentary to the descriptive report/analysis, with Non-parametric clustering analysis for households With sufficient visualization

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
setwd("C:/Users/wb629244/PSSN/pssn2livelihood/03_Endline/03_Endline_HFCs/02_HFC_check_Gdrive_R_prototype")

# Load necessary libraries
library(tidyverse)
library(haven)
library(lubridate)
library(dplyr)
library(tidyr)
library(ggplot2)
library(cluster)
library(factoextra)

source("global_env_setup.R")
source("HFC_utils.R")
```

## Clustering Analysis
The non-parametric analysis below using PAM: Partitioning Around Medoids to discover similarities between households in the PSSN-II baseline dataset. PAM instead of the go to K-means clustering is used, given the prior of the outliers and substantial variance in many variables of interests, e.g. income. 

```{r data_loading_cleaning, echo=FALSE}
hh_constructed = read_dta(file.path(Sys.getenv("analysis_baseline_data_dir"), "HHSurvey.dta"))
hh_numeric <- hh_constructed %>% select(where(is.numeric))

# Filter out columns with too many missing values (e.g. >20%)
threshold = 0.05
hh_filtered <- hh_numeric  %>%
  select(where(~ mean(is.na(.)) <= threshold))


# Filter out constant columns
hh_filtered <- hh_filtered %>%
  select(where(~ var(.) != 0))
```


### Silhouette Analysis for optimal clusters 
```{r optimal_clusters, echo=FALSE}
#Some optimality analysis for # of clusters: a.k.a. silhouette
#Give the size of the dataset, sample only 20% of these households
sample_pct = 0.1
hh_sampled <- hh_filtered %>% sample_frac(sample_pct)
fviz_nbclust(hh_sampled, pam, method ="silhouette")+theme_minimal()
#pam(hh_constructed, k,metric =“manhattan”,stand =FALSE)
```


### Clustering with PAM
```{r clustering, echo=FALSE}
hh_pam_results = pam(hh_sampled,
    k = 3,
    metric = "euclidean",
    stand = FALSE)
fviz_cluster(hh_pam_results, 
             #palette = hh_pam_results$clustering,
             ellipse.type ="euclid",
             repel =TRUE,
             ggtheme =theme_minimal(),
             geom = "point")
```

